---
title: 'Exercise: Reproducibility - Group 1'
author: 'Michael Schulte-Mecklenbeck (Source: https://github.com/datacarpentry)'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  fig_height: 3
  fig_width: 6
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, error = TRUE)
```

This RMarkdown file contains R code you can use to complete the exercise on reproducibility. To see the output simply click on *Knit to HTML* above. This might prompt you to install and load some required 
packages, specifically `knitr`. 

## Packages:

We will use packages from **tidyverse**, a collection of R packages that 
share common philosophies and are designed to work together. 

If you have not yet done so, install tidyverse via `install.packages(tidyverse)`.

Now we need to explicitly load the **tidyverse** packages:

```{r package, message=FALSE, cache=FALSE}
  library("tidyverse")
```

## Part 1: Analyze + document

#### 1. Load the data: 

Note that `read_csv` uses `stringsAsFactors = FALSE` by default.

```{r}
  gap_5060 <- read_csv("../0_data/gapminder-5060.csv")
```

#### 2. Visualize life expectancy over time for Canada in the 1950s and 1960s using a line plot.

```{r}
# Filter the data for Canada only
  gap_5060_CA <- 
    gap_5060 %>%
    filter(country == "Canada")
# Visualize
  ggplot(data = gap_5060_CA, aes(x = year, y = lifeExp)) +
    geom_line()
```

Something is clearly wrong with this plot! Some people live exceptionally long lives.

#### 3. Make appropriate corrections.

Turns out there's a data error in the data file: life expectancy for 
Canada in the year 1957 is coded as `999999`, it should actually be 
`69.96`. We now make this correction:
  
```{r}
  gap_5060 <- 
    gap_5060 %>%
    mutate(lifeExp = replace(lifeExp, (country == "Canada" & year == 1957), 69.96))
```

Hey - we changed the problematic value AND documented it in one go - awesome!
  
#### 4. Visualize life expectancy over time for Canada again, with the corrected data.
  
Exact same code as before, but note that the contents of `gap_5060` 
are different as it has been updated in the previous task.

```{r}
  gap_5060_CA <- 
    gap_5060 %>%
    filter(country == "Canada")

  ggplot(data = gap_5060_CA, aes(x = year, y = lifeExp)) +
    geom_line()
```

#### 4 - Stretch goal. Add lines for Mexico and United States.

- `%in%` for logical operator testing if a country's name is in 
the list provided
- Same visualization code as before, only difference is the input dataset

```{r}
  gap_5060_multiple <- 
    gap_5060 %>%
    filter(country %in% c("Canada", "Mexico", "United States"))

  ggplot(data = gap_5060_multiple, aes(x = year, y = lifeExp, color = country)) +
    geom_line()
```